### 2022-12-24 analyses.Rmd
adding analyses

---
title: "Larkin Secondary CREP"
author: "Erin M. Buchanan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 15)
```

## Libraries

```{r}
# library(rio)
library(lme4)
library(lmerTest)
library(MuMIn)
# library(plyr)
# library(psych)
# library(tidyr)
library(ggplot2)
library(dplyr)
# library(ggmosaic)
# library(MOTE)
# library(flextable)
# library(RColorBrewer)
```

## Import the Data

The `full_long` dataset includes all participants in long format - wherein each trial of their study is on one row of the dataset. This data has been merged with other relevant variables from the ACREP Turri Project found here: https://github.com/doomlab/CREP_Turri

```{r}
load("full_long.Rdata")
```

## Data Screening

- What, if any, rules for exclusion would you like to use?

[89] "age_exclusion" - Age of majority    
[90] "previous_exclusion" - Previously participated 
[92] "studyans_exclusion" - Didn't answer all the questions correctly, or could use when they only get Darrel wrong
[93] "purpose_exclusion" - Guessed the purpose of the study
[95] "lang_exclusion" - Said their language wasn't good

```{r}
DF <- full_long
```

## Examine Data

```{r}
table(DF$j_expertise, 
      DF$vignette, 
      DF$cond,
      useNA = "ifany")

# only use darrel
DF <- DF %>% 
  filter(vignette == "Darrel") %>% 
  filter(!is.na(cond)) %>% 
  filter(!is.na(j_expertise))

# look at table
table(DF$j_expertise, 
      DF$cond,
      useNA = "ifany")

# look at the DVs
tapply(DF$know_vas, list(DF$j_expertise, DF$cond), mean, na.rm = T)
tapply(DF$know_vas, list(DF$j_expertise, DF$cond), sd, na.rm = T)
tapply(DF$know_vas, list(DF$j_expertise, DF$cond), function (x){sum(!is.na(x))})

tapply(DF$luck_vas, list(DF$j_expertise, DF$cond), mean, na.rm = T)
tapply(DF$luck_vas, list(DF$j_expertise, DF$cond), sd, na.rm = T)
tapply(DF$luck_vas, list(DF$j_expertise, DF$cond), function (x){sum(!is.na(x))})

# the bin is reversed
DF$know_bin <- 2 - DF$know_bin
DF$luck_bin <- 2 - DF$luck_bin

DF %>% 
  group_by(j_expertise, cond, know_bin) %>% 
  count()

DF %>% 
  group_by(j_expertise, cond, luck_bin) %>% 
  count()
```

## Knowledge VAS Analysis

We will control for research lab using a multilevel model. We could also control for language, but lab also accounts for this variable.

```{r}
know.model.1 <- lmer(know_vas ~ (1 | person_code),
                    data = DF, 
                    na.action = "na.omit")

know.model.2 <- lmer(know_vas ~ (1 | person_code) + cond*j_expertise,
                    data = DF, 
                    na.action = "na.omit")

AIC(know.model.1)
AIC(know.model.2) # lower = good

summary(know.model.2)
r.squaredGLMM(know.model.2)
```

```{r}
ggplot(DF, aes(cond, know_vas, color = j_expertise, fill = j_expertise)) + 
  stat_summary(fun = mean,
               geom = "bar",
               position = "dodge") +
  stat_summary(fun.data = mean_cl_normal,
               geom = "errorbar", 
               position = position_dodge(width = 0.90),
               width = .2) +
  xlab("Condition") + 
  ylab("Knowledge Visual Analog Scale") + 
  theme_classic() + 
  scale_fill_discrete(name = "Expertise", labels = c("Naive", "Expert")) + 
  scale_color_discrete(name = "Expertise", labels = c("Naive", "Expert"))
```

## Luck VAS Analysis

```{r}
luck.model.1 <- lmer(luck_vas ~ (1 | person_code),
                    data = DF, 
                    na.action = "na.omit")

luck.model.2 <- lmer(luck_vas ~ (1 | person_code) + cond*j_expertise,
                    data = DF, 
                    na.action = "na.omit")

AIC(luck.model.1)
AIC(luck.model.2) # lower = good

summary(luck.model.2)
r.squaredGLMM(luck.model.2)
```

```{r}
ggplot(DF, aes(cond, luck_vas, color = j_expertise, fill = j_expertise)) + 
  stat_summary(fun = mean,
               geom = "bar",
               position = "dodge") +
  stat_summary(fun.data = mean_cl_normal,
               geom = "errorbar", 
               position = position_dodge(width = 0.90),
               width = .2) +
  xlab("Condition") + 
  ylab("Luck Visual Analog Scale") + 
  theme_classic() + 
  scale_fill_discrete(name = "Expertise", labels = c("Naive", "Expert")) + 
  scale_color_discrete(name = "Expertise", labels = c("Naive", "Expert"))
```

## Knowledge Bin Analysis

```{r}
know.model.1 <- glmer(know_bin ~ (1 | person_code),
                    data = DF, 
                    family = binomial,
                    na.action = "na.omit")

know.model.2 <- glmer(know_bin ~ (1 | person_code) + cond*j_expertise,
                    data = DF, 
                    family = binomial,
                    na.action = "na.omit")

AIC(know.model.1)
AIC(know.model.2) # lower = good

summary(know.model.2)
r.squaredGLMM(know.model.2)
```

```{r}
tknow <- as.data.frame(table(DF$know_bin, DF$cond, DF$j_expertise))

colnames(tknow) <- c("knowledge_attribution", "Condition", "Expertise", "Frequency")
tknow$Expertise <- factor(tknow$Expertise,
                          levels = c("naive", "expert"),
                          labels = c("Naive", "Expert"))

tknow %>% 
  ggplot(aes(fill = knowledge_attribution, y = Frequency, x = Expertise)) +
  labs("Figure X. Rates of Knowledge Attribution by Condition and Expertise", fill="knowledge_attribution") +
  geom_bar(position="fill", stat = "identity") +
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage Respondents") +
  scale_fill_grey(name = "Knowledge Attribution", start = .9, end=0, labels = c("Believes", "Knows")) +
  theme_bw()+theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) + 
  facet_wrap(~Condition)  
```

## Luck Bin Analysis

```{r}
luck.model.1 <- glmer(luck_bin ~ (1 | person_code),
                    data = DF, 
                    family = binomial,
                    na.action = "na.omit")

luck.model.2 <- glmer(luck_bin ~ (1 | person_code) + cond*j_expertise,
                    data = DF, 
                    family = binomial,
                    na.action = "na.omit")

AIC(luck.model.1)
AIC(luck.model.2) # lower = good

summary(luck.model.2)
r.squaredGLMM(luck.model.2)
```

```{r}
tluck <- as.data.frame(table(DF$luck_bin, DF$cond, DF$j_expertise))

colnames(tluck) <- c("luck_attribution", "Condition", "Expertise", "Frequency")
tluck$Expertise <- factor(tluck$Expertise,
                          levels = c("naive", "expert"),
                          labels = c("Naive", "Expert"))

tluck %>% 
  ggplot(aes(fill = luck_attribution, y = Frequency, x = Expertise)) +
  labs("Figure X. Rates of Inability/Luck Attribution by Condition and Expertise", fill="luck_attribution") +
  geom_bar(position="fill", stat = "identity") +
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage Respondents") +
  scale_fill_grey(name = "Inability/Luck Attribution", start = .9, end=0, labels = c("Inability", "Luck")) +
  theme_bw()+theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) + 
  facet_wrap(~Condition)  
```